---
project:
  output: out
title: "CMV Simulation"
format:
  html:
    page-layout: full
    theme: 
      - cosmo 
      - custom.scss
    monofont: "Fira Code"
    df-print: paged
---

## CMV Simulation

This document will go through simulating CMV and IEP.

## Initialization

We will need the following projects:

- [tidyverse](https://tidyverse.org): Data wrangling
- [lavaan](https://lavaan.ugent.be): SEM functionality
- [semPlot](http://sachaepskamp.com/documentation/semPlot/semPaths.html): Plotting SEM diagrams

We'll also set the random seed to an arbitrary fixed value to allow for reproduceability

```{r}
#| output: false
library(tidyverse)
library(lavaan)
library(semPlot)

set.seed(1337)
```

## Model

Our model has three latent variables that are each measured with several items:

- Mobile phone addiction: phone_add
  
  1. Missed Work: mw
  2. Concentration: conc
  3. Pain: pain
  4. Require Phone: reqp
  5. Impatient: imp
  6. Thinking: think
  7. Never Give Up: ngu
  8. Constant Checking: check
  9. Extended Usage: use
  10. Perceived Extended Usage: puse
  
- Perceived Stress: stress

  1. Upset: upset
  2. Unable to Control: nocont
  3. Nervous: nerv
  4. Irritating Hassles: hass
  5. Coping: cope
  6. Confident: conf
  7. Going Your Way: way
  8. No Cope: nocope
  9. Control: cont
  10. On Top: ontop
  11. Angered: angr
  12. Anxious: anx
  13. Control Time: time
  14. Piling: pile
  
- Perceived Learning: plearn

  1. Organize: org
  2. Guide: guid
  3. Use Knowledge: usek
  4. Attitude: att
  5. Critique: crit
  6. Self-Reliant: selfr
  7. Sophisticated: soph
  
We'll input the model in lavaan formula syntax and have it create an SEM diagram for us  

```{r}
#| fig-width: 10
#| fig-height: 10
cfa_model <- '
  # measurement model
  phone_add =~ mw + conc + pain + reqp + imp + think + ngu + check + use + puse
  stress =~ upset + nocont + nerv + hass + cop + conf + way + nocope + cont + ontop + angr + anx + time + pile
  plearn =~ org + guid + usek + att + crit + selfr + soph
  
  # covariances
  phone_add ~~ stress
  phone_add ~~ plearn
  stress ~~ plearn
'

semPlotModel_lavaanModel(cfa_model) |> 
  semPaths(layout="tree", rotation=2, nCharNodes=5, sizeMan2 = 2.5, mar = c(1,4,1,3), label.norm = "OOOOO", residuals = FALSE)
```

## Simulation

Now we'll simulate data for this model.  In this case all item weights are simulated as being equal

::: {.callout-note}
Right now everything is coded in the same direction. Will need to fix loading at -1 for reverse coded items
:::


```{r}
cfa_sim <- simulateData(cfa_model, model.type = "cfa", sample.nobs = 500)
#change to likert scores
cfa_sim <- cfa_sim |> mutate(across(everything(), \(x) findInterval(x, vec=c(-Inf,-2.5,-1, 1,2.5,Inf))))
cfa_sim
```

Check the fit of the simulation

```{r}
fit <- cfa(cfa_model, cfa_sim)
summary(fit, fit.measures = TRUE)
```



